Flowchart of Current Algorithm

### 2008-10-24 docs/flowchart/flowchart.tex
separate out the content so we can produce a master document
\ut -> \Ut to avoid multiple definitions

\documentclass[11pt]{article}

\tolerance=600

\usepackage{amsmath,color,longtable}

% Margins
\usepackage[lmargin=0.5in,rmargin=1.5in,tmargin=1.0in,bmargin=1.0in]{geometry}

\newcommand{\rhozero}{\rho_0}
\newcommand{\pizero}{\pi_0}
\newcommand{\pizeroone}{\pi_0^{(1)}}
\newcommand{\pizerotwo}{\pi_0^{(2)}}
\newcommand{\gammabar}{\overline{\Gamma}_1}

\newcommand{\nablab}{\mathbf{\nabla}}
\newcommand{\cdotb}{\mathbf{\cdot}}

\newcommand{\sfrac}[2]{\mathchoice
  {\kern0em\raise.5ex\hbox{\the\scriptfont0 #1}\kern-.15em/
   \kern-.15em\lower.25ex\hbox{\the\scriptfont0 #2}}
  {\kern0em\raise.5ex\hbox{\the\scriptfont0 #1}\kern-.15em/
   \kern-.15em\lower.25ex\hbox{\the\scriptfont0 #2}}
  {\kern0em\raise.5ex\hbox{\the\scriptscriptfont0 #1}\kern-.2em/
   \kern-.15em\lower.25ex\hbox{\the\scriptscriptfont0 #2}}
  {#1\!/#2}}

\newcommand{\half}{\sfrac{1}{2}}
\newcommand{\myhalf}{\sfrac{1}{2}}
\newcommand{\nph}{{n + \myhalf}}
\newcommand{\nmh}{{n - \myhalf}}

\newcommand{\Hext}{{H_{\rm ext}}}
\newcommand{\Hnuc}{{H_{\rm nuc}}}
\newcommand{\kth}{k_{\rm th}}

\newcommand{\Sbar}{\overline{S}}

\newcommand{\inp}{\mathrm{in}}
\newcommand{\initp}{\mathrm{init}}
\newcommand{\outp}{\mathrm{out}}

\newcommand{\uadv}{\widetilde{\mathbf{U}}^{\mathrm{ADV}}}
\newcommand{\uadvone}{\widetilde{\mathbf{U}}^{\mathrm{ADV},\star}}
\newcommand{\uadvtwo}{\widetilde{\mathbf{U}}^{\mathrm{ADV}}}
\newcommand{\V}{\mathbf{V}}
\newcommand{\dt}{\Delta t}
\newcommand{\dr}{\Delta r}

\newcommand{\etarho}{\eta_{\rho}}

\newcommand{\ubold}{\mathbf{U}}
\newcommand{\ut}{\widetilde{\ubold}}

\newcommand{\omegadot}{\dot{\omega}}
\newcommand{\er}{\mathbf{e}_r}

\setlength{\marginparwidth}{1.0in}
\newcommand{\MarginPar}[1]{\marginpar{%
\vskip-\baselineskip %raise the marginpar a bit
\raggedright\tiny\sffamily
\hrule\smallskip{\color{red}#1}\par\smallskip\hrule}}

\title{Flowchart of Current Algorithm}

\begin{document}

\maketitle
\tableofcontents
\cleardoublepage

@@

$c_p$   & specific heat at constant pressure
          ($c_p \equiv \left . \partial h / \partial T \right |_{p,X_k}$)
        & erg~g$^{-1}$~K$^{-1}$ \\
\hline
$f$     & volume discrepancy factor ($0 \le f \le 1$) & -- \\
\hline
$g$     & gravitational acceleration                 & cm~s$^{-2}$ \\
\hline
$h$     & specific enthalpy                          & erg~g$^{-1}$ \\
\hline
$\Hext$ & external heating energy generation rate    & erg~g$^{-1}$~s$^{-1}$ \\
\hline
$\Hnuc$ & nuclear energy generation rate             & erg~g$^{-1}$~s$^{-1}$ \\
\hline
$h_p$   & $h_p \equiv \left . \partial h / \partial p \right |_{T,X_k}$ & cm$^{3}$~g$^{-1}$ \\
\hline
$\kth$  & thermal conductivity                       & erg~cm$^{-1}$~s$^{-1}$~K$^{-1}$ \\
\hline
$p_0$   & base state pressure                        & erg~cm$^{-3}$ \\
\hline
$p_T$   & $p_T \equiv \left . \partial p / \partial T \right |_{\rho,X_k}$ & erg~cm$^{-3}$~K$^{-1}$ \\
\hline
$p_{X_k}$ & $p_{X_k} \equiv \left . \partial p / \partial X_k \right |_{p,T,X_{j,j\ne k}}$ & erg~cm$^{-3}$ \\
\hline
$p_\rho$ & $p_\rho \equiv \left . \partial p / \partial \rho \right |_{T,X_k}$ & erg~g$^{-1}$ \\
\hline
$q_k$   & specific nuclear binding energy            & erg~g$^{-1}$  \\
\hline
$r$     & radial coordinate (direction of gravity)   & cm \\
\hline
$s$     & specific entropy                           & erg~g$^{-1}$~K$^{-1}$ \\
\hline
$S$     & source term to the divergence constraint   & s$^{-1}$ \\
\hline
$t$     & time                                       & s \\
\hline
$T$     & temperature                                & K \\
\hline
$\ubold$     & total velocity ($\ubold = \ut + w_0 \er$) & cm~s$^{-1}$ \\
\hline
$\ut$   & local velocity                             & cm~s$^{-1}$ \\
\hline
$\uadv$ & advective velocity (edge-centered)         & cm~s$^{-1}$ \\
\hline
$w_0$   & base state expansion velocity              & cm~s$^{-1}$ \\
\hline
$X_k$   & mass fraction of the species ($\sum_k X_k = 1$) & -- \\
\hline
$\beta_0$ & coefficient to velocity
            in velocity constraint equation  & g~cm$^{-3}$ \\
\hline
$\Gamma_1$ & first adiabatic exponent ($\Gamma_1 \equiv \left . d \log p/d \log \rho \right |_s$) & -- \\
\hline
$\etarho$ & $\etarho \equiv \overline{ (\rho' \ut \cdot \er)}$ & g~cm$^{-2}$~s$^{-1}$ \\
\hline
$\xi_k$ & $\xi_k \equiv \left . \partial h / \partial X_k \right |_{p,T,X_{j,j\ne k}}$ & erg~g$^{-1}$ \\
\hline 
$\pi$   & dynamic pressure & erg~cm$^{-3}$ \\
\hline
$\pizero$ & base state dynamic pressure & erg~cm$^{-3}$ \\
\hline
$\rho$  & mass density  & g~cm$^{-3}$ \\
\hline
$\rho_0$  & base state mass density  & g~cm$^{-3}$ \\
\hline
$\rho'$  & perturbational density ($\rho' = \rho - \rho_0$) & g~cm$^{-3}$ \\
\hline
$(\rho h)_0$ & base state enthalpy density & erg~cm$^{-3}$  \\
\hline
$(\rho h)'$ & perturbational enthalpy density 
              $ \left [(\rho h)' = \rho h - (\rho h)_0 \right ]$ & erg~cm$^{-3}$  \\
\hline
$\sigma$ & $\sigma \equiv p_T/(\rho c_p p_\rho)$ & erg$^{-1}$~g \\
\hline
$\psi$  & $\psi \equiv D_0 p_0/Dt$ & erg~cm$^{-3}$~s$^{-1}$ \\
\hline
$\omegadot_k$ & creation rate for species $k$ ($\omegadot_k \equiv DX_k/Dt$) & s$^{-1}$ \\
\end{longtable}
\end{center}
%
\renewcommand{\arraystretch}{1.0}

@@

The procedure to construct $\uadvone$ is described in detail in Appendix B of paper III.
We note that  $\uadvone$ satisfies the discrete versions of 
\begin{equation}
\int_{\Omega_H} \ut \cdotb \er \; dA = 0 \label{eq:udoterzero} \enskip ,
\end{equation}
(paper III; eq.~[21])
and
\begin{equation}
\nablab \cdotb (\beta_0 \ut )  = \beta_0 \left(S - \Sbar \right )\enskip ,
\label{eq:tildeconstraint}
\end{equation}
(paper III; eq.~[39]).  The discretization with the volume discrepancy correction is
\begin{equation}
\nablab \cdotb \left(\beta_0^n \uadvone\right) = 
\beta_0^n \left(S^{\nph,\star} - \overline{S}^{\nph,\star}\right)
+ \frac{f}{\gammabar^n p_0^n}
\left[\frac{p(\rho,h,X_k)^n - \overline{p(\rho,h,X_k)^n}}{\Delta t^n}\right] \enskip .
\end{equation}

@@

The velocity update happens analogously to paper II, using $S^{n+1}$ from step 10.
We update the velocity field $\ut^n$ to $\ut^{n+1,\star}$ by discretizing
\begin{equation}
\frac{\partial\ut}{\partial t} = - \ut \cdotb \nablab \ut - w_0 \frac{\partial \ut}{\partial r}
                                 - \left(\ut \cdotb \er\right) \frac{\partial w_0}{\partial r} \er
                                 - \frac{1}{\rho} \nablab\pi
                                 + \frac{1}{\rho_0} \frac{\partial \pizero}{\partial r} \er
                                 - \frac{(\rho-\rhozero)}{\rho} \; g \; \er  \label{eq:utildeupd}  \enskip\
 ,
\end{equation}
(paper III; eq.~[37]) as 
\begin{eqnarray}
\ut^{n+1,\star} &=& \ut^n - \dt \;
 \left[\left(\uadvtwo+ w_0^{\nph} \er\right) \cdotb \nablab \ut \right]^\nph
      - \dt \; \left(\uadvtwo \cdotb \er\right)  \left(\frac{\partial w_0^{\nph}}{\partial r} \right) \er \nonumber \\
   &&   + \dt \left[ - \frac{1}{\rho^\nph} {\mathbf{G}} \pi^\nmh
        + \frac{1}{\rho_0} {\mathbf{G}} \pizerotwo
        - \frac{(\rho^\nph-\rhozero^\nph)}{\rho^\nph} g \er \right] \enskip ,
\end{eqnarray}
where $\rho^\nph = ( \rho^n + \rho^{n+1} )/2$
and $\mathbf{G}$ approximates a cell-centered gradient from nodal data. 
The construction of $[(\uadvtwo+ w_0^{\nph} \er \cdotb \nablab) \ut ]^\nph$
is described in paper II, Appendix A, with $\V = \uadvtwo+w_0^{\nph} \er$ and~$s$ set to each
component of $\ut^n$ individually.

Finally, we impose the divergence constraint from equation (\ref{eq:tildeconstraint}) 
including the volume discrepancy correction,
\begin{equation}
\nablab \cdotb \left(\beta_0^{\nph} \ut^{n+1} \right)  = \beta_0^{\nph} \left(S^{n+1} - \overline{S}^{n+1} \right)
+ \frac{f}{\gammabar^{n+1} p_0^{n+1}}
\left[\frac{p(\rho,h,X_k)^{n+1} - \overline{p(\rho,h,X_k)^{n+1}}}{\Delta t^n}\right]
\enskip,
\end{equation}
by defining ${\bf V} = \ut^{n+1,\star} + ( \dt / \rho^\nph ) \; \mathbf{G} \pi^\nmh$ 
and solving
\begin{equation}
 L_\beta^\rho \phi =
   D \left ( \beta_0^{\nph} {\bf V} \right) - \beta_0^{\nph} \left(S^{n+1}-\overline{S}^{n+1} \right)
- \frac{f}{\gammabar^{n+1} p_0^{n+1}}
\left[\frac{p(\rho,h,X_k)^{n+1} - \overline{p(\rho,h,X_k)^{n+1}}}{\Delta t^n}\right] \enskip,
\end{equation}
for nodal values of $\phi$, where
$\beta_0^\nph = ( \beta_0^n + \beta_0^{n+1} )/2$ and
$L_\beta^\rho$ is the standard bilinear
finite element approximation to $\nablab \cdotb ({\beta_0^\nph}/{\rho^\nph}) \nablab.$
\MarginPar{Why are $\beta$ and $\rho$ time centered here instead using the new time value?  Does this have something to do with how $\phi$ relates to a lagged pressure?}
In this step, $D$ is a discrete second-order operator that approximates the 
divergence at nodes from cell-centered data and satisfies
$\mathbf{G} = -D^T.$ 
(See (cite almgrenBellSzymczak:1996) for a detailed discussion of this
approximate projection; see (cite almgren:bell:crutchfield) for a discussion
of this particular form of the projection operand.)  
We solve the linear system of equation
\begin{equation}
 L_\beta^\rho \phi =
   D \left ( \beta_0^{\nph} {\bf V} \right) - \beta_0^{\nph} \left(S^{n+1}-\overline{S}^{n+1} \right)\enskip,
\label{eq:nodal_solve}
\end{equation}
(paper III; eq.~[77]) using multigrid V-cycles with Gauss-Seidel relaxation.

We determine the new time velocity field from
\begin{equation}
\ut^{n+1} = {\bf V} - \frac{1}{\rho^\nph} \mathbf{G} \phi \enskip ,
\end{equation}
and the new time-centered perturbational pressure from
\begin{equation}
  \pi^\nph = \frac{1}{\dt} \phi \enskip .
\end{equation}

@@

Compute $\dt$ for the next time step with the procedure described in \S~ref{sec:timesteps},
using $w_0$ as computed in step 6 and
$\ut^{n+1}$ as computed in step 11.  We use this $\dt$ in the next time step.

@@

  \begin{enumerate}
  \item Perform steps 1--11 as described above, using 
    $S^{\half,\star} = (S^0 + S^{1,\nu})/2$ in step 1 as described.
    The only other difference in the time advancement is that in step 11
    we define ${\bf V} = (\ut^{1,\star} - \ut^0)$ and solve
    \begin{equation}  L_\beta^\rho \phi =
      D \left ( \beta_0^{\half} {\bf V} \right) - \beta_0^{\half} \left[ (S^{1}-\overline{S}^{1}) - (S^{0}-\overline{S}^{0}) \right] \enskip . 
    \end{equation}
    (The motivation for this form of the projection in the initial pressure iterations
    is discussed in (cite almgren:bell:crutchfield).)
      We discard the new velocity resulting from this, but keep the new  
      value for $\pi^{\half} = \pi^{-\half} + (1 / \dt) \; \phi.$  
      These steps also yield new scalar data at time $\dt,$ which
      we discard,  and new values for $\etarho^{\half}$ (step 8), $\psi^{\half}$ (step 8), 
      $S^{1,\nu+1}$ (step 10), and $\pi^{\half}$ (step 11), which we keep.
    \item Set $\pi^{-\half} = \pi^{\half}$, $\etarho^{-\half} = \etarho^{\half}$,
      and $\psi^{-\half} = \psi^{\half}$. 
    \end{enumerate}

@@

\end{document}

>>>>

@@

$c_p$   & specific heat at constant pressure
          ($c_p \equiv \left . \partial h / \partial T \right |_{p,X_k}$)
        & erg~g$^{-1}$~K$^{-1}$ \\
\hline
$f$     & volume discrepancy factor ($0 \le f \le 1$) & -- \\
\hline
$g$     & gravitational acceleration                 & cm~s$^{-2}$ \\
\hline
$h$     & specific enthalpy                          & erg~g$^{-1}$ \\
\hline
$\Hext$ & external heating energy generation rate    & erg~g$^{-1}$~s$^{-1}$ \\
\hline
$\Hnuc$ & nuclear energy generation rate             & erg~g$^{-1}$~s$^{-1}$ \\
\hline
$h_p$   & $h_p \equiv \left . \partial h / \partial p \right |_{T,X_k}$ & cm$^{3}$~g$^{-1}$ \\
\hline
$\kth$  & thermal conductivity                       & erg~cm$^{-1}$~s$^{-1}$~K$^{-1}$ \\
\hline
$p_0$   & base state pressure                        & erg~cm$^{-3}$ \\
\hline
$p_T$   & $p_T \equiv \left . \partial p / \partial T \right |_{\rho,X_k}$ & erg~cm$^{-3}$~K$^{-1}$ \\
\hline
$p_{X_k}$ & $p_{X_k} \equiv \left . \partial p / \partial X_k \right |_{p,T,X_{j,j\ne k}}$ & erg~cm$^{-3}$ \\
\hline
$p_\rho$ & $p_\rho \equiv \left . \partial p / \partial \rho \right |_{T,X_k}$ & erg~g$^{-1}$ \\
\hline
$q_k$   & specific nuclear binding energy            & erg~g$^{-1}$  \\
\hline
$r$     & radial coordinate (direction of gravity)   & cm \\
\hline
$s$     & specific entropy                           & erg~g$^{-1}$~K$^{-1}$ \\
\hline
$S$     & source term to the divergence constraint   & s$^{-1}$ \\
\hline
$t$     & time                                       & s \\
\hline
$T$     & temperature                                & K \\
\hline
$\ubold$     & total velocity ($\ubold = \Ut + w_0 \er$) & cm~s$^{-1}$ \\
\hline
$\Ut$   & local velocity                             & cm~s$^{-1}$ \\
\hline
$\uadv$ & advective velocity (edge-centered)         & cm~s$^{-1}$ \\
\hline
$w_0$   & base state expansion velocity              & cm~s$^{-1}$ \\
\hline
$X_k$   & mass fraction of the species ($\sum_k X_k = 1$) & -- \\
\hline
$\beta_0$ & coefficient to velocity
            in velocity constraint equation  & g~cm$^{-3}$ \\
\hline
$\Gamma_1$ & first adiabatic exponent ($\Gamma_1 \equiv \left . d \log p/d \log \rho \right |_s$) & -- \\
\hline
$\etarho$ & $\etarho \equiv \overline{ (\rho' \Ut \cdot \er)}$ & g~cm$^{-2}$~s$^{-1}$ \\
\hline
$\xi_k$ & $\xi_k \equiv \left . \partial h / \partial X_k \right |_{p,T,X_{j,j\ne k}}$ & erg~g$^{-1}$ \\
\hline 
$\pi$   & dynamic pressure & erg~cm$^{-3}$ \\
\hline
$\pizero$ & base state dynamic pressure & erg~cm$^{-3}$ \\
\hline
$\rho$  & mass density  & g~cm$^{-3}$ \\
\hline
$\rho_0$  & base state mass density  & g~cm$^{-3}$ \\
\hline
$\rho'$  & perturbational density ($\rho' = \rho - \rho_0$) & g~cm$^{-3}$ \\
\hline
$(\rho h)_0$ & base state enthalpy density & erg~cm$^{-3}$  \\
\hline
$(\rho h)'$ & perturbational enthalpy density 
              $ \left [(\rho h)' = \rho h - (\rho h)_0 \right ]$ & erg~cm$^{-3}$  \\
\hline
$\sigma$ & $\sigma \equiv p_T/(\rho c_p p_\rho)$ & erg$^{-1}$~g \\
\hline
$\psi$  & $\psi \equiv D_0 p_0/Dt$ & erg~cm$^{-3}$~s$^{-1}$ \\
\hline
$\omegadot_k$ & creation rate for species $k$ ($\omegadot_k \equiv DX_k/Dt$) & s$^{-1}$ \\
\end{longtable}
\end{center}
%
\renewcommand{\arraystretch}{1.0}

@@

The procedure to construct $\uadvone$ is described in detail in Appendix B of paper III.
We note that  $\uadvone$ satisfies the discrete versions of 
\begin{equation}
\int_{\Omega_H} \Ut \cdotb \er \; dA = 0 \label{eq:udoterzero} \enskip ,
\end{equation}
(paper III; eq.~[21])
and
\begin{equation}
\nablab \cdotb (\beta_0 \Ut )  = \beta_0 \left(S - \Sbar \right )\enskip ,
\label{eq:tildeconstraint}
\end{equation}
(paper III; eq.~[39]).  The discretization with the volume discrepancy correction is
\begin{equation}
\nablab \cdotb \left(\beta_0^n \uadvone\right) = 
\beta_0^n \left(S^{\nph,\star} - \overline{S}^{\nph,\star}\right)
+ \frac{f}{\gammabar^n p_0^n}
\left[\frac{p(\rho,h,X_k)^n - \overline{p(\rho,h,X_k)^n}}{\Delta t^n}\right] \enskip .
\end{equation}

@@

The velocity update happens analogously to paper II, using $S^{n+1}$ from step 10.
We update the velocity field $\Ut^n$ to $\Ut^{n+1,\star}$ by discretizing
\begin{equation}
\frac{\partial\Ut}{\partial t} = - \Ut \cdotb \nablab \Ut - w_0 \frac{\partial \Ut}{\partial r}
                                 - \left(\Ut \cdotb \er\right) \frac{\partial w_0}{\partial r} \er
                                 - \frac{1}{\rho} \nablab\pi
                                 + \frac{1}{\rho_0} \frac{\partial \pizero}{\partial r} \er
                                 - \frac{(\rho-\rhozero)}{\rho} \; g \; \er  \label{eq:utildeupd}  \enskip\
 ,
\end{equation}
(paper III; eq.~[37]) as 
\begin{eqnarray}
\Ut^{n+1,\star} &=& \Ut^n - \dt \;
 \left[\left(\uadvtwo+ w_0^{\nph} \er\right) \cdotb \nablab \Ut \right]^\nph
      - \dt \; \left(\uadvtwo \cdotb \er\right)  \left(\frac{\partial w_0^{\nph}}{\partial r} \right) \er \nonumber \\
   &&   + \dt \left[ - \frac{1}{\rho^\nph} {\mathbf{G}} \pi^\nmh
        + \frac{1}{\rho_0} {\mathbf{G}} \pizerotwo
        - \frac{(\rho^\nph-\rhozero^\nph)}{\rho^\nph} g \er \right] \enskip ,
\end{eqnarray}
where $\rho^\nph = ( \rho^n + \rho^{n+1} )/2$
and $\mathbf{G}$ approximates a cell-centered gradient from nodal data. 
The construction of $[(\uadvtwo+ w_0^{\nph} \er \cdotb \nablab) \Ut ]^\nph$
is described in paper II, Appendix A, with $\V = \uadvtwo+w_0^{\nph} \er$ and~$s$ set to each
component of $\Ut^n$ individually.

Finally, we impose the divergence constraint from equation (\ref{eq:tildeconstraint}) 
including the volume discrepancy correction,
\begin{equation}
\nablab \cdotb \left(\beta_0^{\nph} \Ut^{n+1} \right)  = \beta_0^{\nph} \left(S^{n+1} - \overline{S}^{n+1} \right)
+ \frac{f}{\gammabar^{n+1} p_0^{n+1}}
\left[\frac{p(\rho,h,X_k)^{n+1} - \overline{p(\rho,h,X_k)^{n+1}}}{\Delta t^n}\right]
\enskip,
\end{equation}
by defining ${\bf V} = \Ut^{n+1,\star} + ( \dt / \rho^\nph ) \; \mathbf{G} \pi^\nmh$ 
and solving
\begin{equation}
 L_\beta^\rho \phi =
   D \left ( \beta_0^{\nph} {\bf V} \right) - \beta_0^{\nph} \left(S^{n+1}-\overline{S}^{n+1} \right)
- \frac{f}{\gammabar^{n+1} p_0^{n+1}}
\left[\frac{p(\rho,h,X_k)^{n+1} - \overline{p(\rho,h,X_k)^{n+1}}}{\Delta t^n}\right] \enskip,
\end{equation}
for nodal values of $\phi$, where
$\beta_0^\nph = ( \beta_0^n + \beta_0^{n+1} )/2$ and
$L_\beta^\rho$ is the standard bilinear
finite element approximation to $\nablab \cdotb ({\beta_0^\nph}/{\rho^\nph}) \nablab.$
\MarginPar{Why are $\beta$ and $\rho$ time centered here instead using the new time value?  Does this have something to do with how $\phi$ relates to a lagged pressure?}
In this step, $D$ is a discrete second-order operator that approximates the 
divergence at nodes from cell-centered data and satisfies
$\mathbf{G} = -D^T.$ 
(See (cite almgrenBellSzymczak:1996) for a detailed discussion of this
approximate projection; see (cite almgren:bell:crutchfield) for a discussion
of this particular form of the projection operand.)  
We solve the linear system of equation
\begin{equation}
 L_\beta^\rho \phi =
   D \left ( \beta_0^{\nph} {\bf V} \right) - \beta_0^{\nph} \left(S^{n+1}-\overline{S}^{n+1} \right)\enskip,
\label{eq:nodal_solve}
\end{equation}
(paper III; eq.~[77]) using multigrid V-cycles with Gauss-Seidel relaxation.

We determine the new time velocity field from
\begin{equation}
\Ut^{n+1} = {\bf V} - \frac{1}{\rho^\nph} \mathbf{G} \phi \enskip ,
\end{equation}
and the new time-centered perturbational pressure from
\begin{equation}
  \pi^\nph = \frac{1}{\dt} \phi \enskip .
\end{equation}

@@

Compute $\dt$ for the next time step with the procedure described in \S~ref{sec:timesteps},
using $w_0$ as computed in step 6 and
$\Ut^{n+1}$ as computed in step 11.  We use this $\dt$ in the next time step.

@@

  \begin{enumerate}
  \item Perform steps 1--11 as described above, using 
    $S^{\half,\star} = (S^0 + S^{1,\nu})/2$ in step 1 as described.
    The only other difference in the time advancement is that in step 11
    we define ${\bf V} = (\Ut^{1,\star} - \Ut^0)$ and solve
    \begin{equation}  L_\beta^\rho \phi =
      D \left ( \beta_0^{\half} {\bf V} \right) - \beta_0^{\half} \left[ (S^{1}-\overline{S}^{1}) - (S^{0}-\overline{S}^{0}) \right] \enskip . 
    \end{equation}
    (The motivation for this form of the projection in the initial pressure iterations
    is discussed in (cite almgren:bell:crutchfield).)
      We discard the new velocity resulting from this, but keep the new  
      value for $\pi^{\half} = \pi^{-\half} + (1 / \dt) \; \phi.$  
      These steps also yield new scalar data at time $\dt,$ which
      we discard,  and new values for $\etarho^{\half}$ (step 8), $\psi^{\half}$ (step 8), 
      $S^{1,\nu+1}$ (step 10), and $\pi^{\half}$ (step 11), which we keep.
    \item Set $\pi^{-\half} = \pi^{\half}$, $\etarho^{-\half} = \etarho^{\half}$,
      and $\psi^{-\half} = \psi^{\half}$. 
    \end{enumerate}

@@
